---
---
title: "Bellabeat Smart Device Data Analysis"
author: "Data Analyst"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(scales)
library(kableExtra)
```

## Introduction
This report analyzes data collected from FitBit users to generate insights for Bellabeat, a wellness technology company. The objective is to understand user behavior and usage patterns to inform future product and marketing strategies.

## Data Loading
```{r load-data}
activity <- read_csv("Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
calories <- read_csv("Fitabase Data 4.12.16-5.12.16/hourlyCalories_merged.csv")
steps <- read_csv("Fitabase Data 4.12.16-5.12.16/hourlySteps_merged.csv")
sleep <- read_csv("Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")
```

## Data Cleaning
```{r clean-data}
activity <- activity %>% mutate(ActivityDate = mdy(ActivityDate))
sleep <- sleep %>% mutate(SleepDay = mdy_hms(SleepDay)) %>% 
  group_by(Id, date = as.Date(SleepDay)) %>% 
  summarise(TotalSleep = sum(TotalMinutesAsleep), .groups = 'drop')
```

## Summary Statistics
```{r summary-stats}
activity %>% 
  summarise(across(where(is.numeric), list(mean = mean, median = median), na.rm = TRUE)) %>% 
  kable() %>% 
  kable_styling()
```

## Visualizations
### Steps Distribution
```{r steps-plot}
activity %>% 
  ggplot(aes(TotalSteps)) + 
  geom_histogram(binwidth = 1000, fill = "skyblue", color = "black") + 
  labs(title = "Distribution of Daily Steps", x = "Steps", y = "Frequency")
```

### Calories vs Steps
```{r calories-vs-steps}
activity %>% 
  ggplot(aes(TotalSteps, Calories)) + 
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm") +
  labs(title = "Calories Burned vs. Steps Taken", x = "Total Steps", y = "Calories")
```

### Sleep vs Activity
```{r sleep-vs-activity}
combined <- inner_join(activity, sleep, by = c("Id" = "Id", "ActivityDate" = "date"))
combined %>% 
  ggplot(aes(TotalSleep, TotalSteps)) + 
  geom_point(alpha = 0.6) + 
  labs(title = "Sleep Duration vs. Steps Taken", x = "Sleep (minutes)", y = "Steps")
```

## Insights
- Most users average fewer than 8,000 steps per day.
- There is a positive correlation between steps taken and calories burned.
- Users with longer sleep durations tend to show slightly higher physical activity.

## Recommendations
1. Promote daily step goals with app notifications.
2. Emphasize the link between sleep and activity in wellness tips.
3. Use behavior clustering to tailor in-app experiences.

## Conclusion
This analysis provides a baseline understanding of user health behaviors, which Bellabeat can leverage to refine products and drive user engagement.
